Remote Collection of Respiratory Samples

### 2024-07-12 Jordana Hoppe/Home Collection/manuscript.qmd
figure updates for Jordana

---
title: "Remote Collection of Respiratory Samples"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
editor: source
---

```{r libraries}
#| include: false
library(tidyverse)
library(lubridate)
library(gtsummary)
library(irr)
library(eulerr)
library(kableExtra)
library(caret)
library(likert)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning, include=FALSE}
# Use REDCap import
source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_data.r")
source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r")
# Exclude missing subjects
data_exclude <- c("009-07", "009-17", "009-25", "009-37")
survey_exclude <- c(15, 27, 30, 34, 47)
# Fix missing labels
labels <- label(data)[which(label(data) != "")]
labels[paste0(names(labels), ".factor")] <- as.character(labels)
survey_labels <- label(survey_data)[which(label(survey_data) != "")]
survey_labels[paste0(names(survey_labels), ".factor")] <- as.character(survey_labels)
labels <- c(labels, survey_labels)
# Combine race columns
races <- c(
  "American Indian or Alaska Native", "Asian", "Black or African-American",
  "Native Hawaiian or Other Pacific Islander", "White", "More than one race",
  "Unknown or not reported"
)
data$race <- apply(data, 1, function(r) {
  w <- which(r[paste0("race___", 1:7)] == 1)
  paste0(races[w], collapse = "/")
})
# Age
data$age <- round(as.numeric(ymd(data$clinic_date) - ymd(data$pt_dob)) / 365.25, 1)
# Find tests with no growth
data$clinic_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("clinic_cx_results___\\d$", names(r))]), na.rm = T)
})
data$clinic_cx_no_growth.factor <- factor(data$clinic_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
data$res_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("res_cx_results___\\d$", names(r))]), na.rm = T)
})
data$res_cx_no_growth.factor <- factor(data$res_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
data$home_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("home_cx_results___\\d$", names(r))]), na.rm = T)
})
data$home_cx_no_growth.factor <- factor(data$home_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
# Additional labels for tables
labels["race"] <- "Race"
labels["age"] <- "Age at Clinic Visit"
# Genotype format
data$Genotype <- paste0(
  data$mutation1.factor == "F508del", "/", data$mutation2.factor == "F508del"
)
data$Genotype <- factor(data$Genotype,
  levels = c("TRUE/TRUE", "TRUE/FALSE", "FALSE/FALSE"),
  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
```

@@

## Table 3: Clinic Culture Results

```{r}
#| label: tbl-3
#| tbl-cap: Clinic Culture Results
data %>%
  filter(!subject_id %in% data_exclude) %>%
  select(clinic_cx_source.factor:clinic_burk_sp.factor) %>%
  tbl_summary(
    label = labels, missing_text = "Missing"
  ) %>%
  as_kable_extra()
```

@@

```{r}
agreement <- function(result_pattern, df = data) {
  cols <- paste0(
    c("clinic_cx_", "res_cx_", "home_cx_"),
    paste0(result_pattern, ".factor")
  )
  d <- df[, cols]
  colnames(d) <- c("Clinic", "Delayed", "Home")
  d <- d %>%
    mutate(across(everything(), ~ as.character(.)))
  d[d == "NA"] <- NA
  d[d == "Unchecked"] <- "Absent"
  d[d == "Checked"] <- "Present"
  fk <- kappam.fleiss(d)
  cat("\n")
  cat("### Agreement between all three methods")
  cat("\n")
  cat("\n")
  cat("#### Fleiss' kappa")
  cat("\n")
  cat("\n")
  f <- ftable(d)
  f %>%
    as.data.frame(.) %>%
    kable(.) %>%
    kable_styling(., "striped") %>%
    print(.)
  cat("\n")
  cat("\n")
  cat(
    "Overall Fleiss' kappa for the three tests was", round(fk$value, 3),
    paste0("(p = ", format.pval(fk$p.value, digits = 3, eps = 0.001), ")")
  )
  cat("\n")
  cat("\n")
  cat("#### Euler diagram")
  cat("\n")
  d %>%
    mutate(across(everything(), ~ as.numeric(. == "Present" | . == "No Growth"))) %>%
    drop_na(.) %>%
    euler(.) %>%
    plot(., quantities = T) %>%
    print(.)
  cat("\n")
  cat("\n")
  cat("### Pairwise agreement")
  cat("\n")
  d <- data.frame(lapply(d, factor, levels = c("Absent", "Present")))
  cat("\n")
  cat("#### Clinic vs. delayed (clinic as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Delayed"]),
    reference = d[, "Clinic"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("#### Clinic vs. home (clinic as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Home"]),
    reference = d[, "Clinic"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("#### Home vs. delayed (delayed as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Home"]),
    reference = d[, "Delayed"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
}
```

@@

```{r}
yn_df <- t %>%
  select(all_of(yn)) %>%
  rename(any_of(setNames(names(labels), labels)))
yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")), check.names = F)
plot(likert(yn_df))
```

```{r}
check_df <- t %>%
  select(all_of(checkboxes)) %>%
  rename(any_of(setNames(names(labels), labels)))
plot(likert(check_df))
```

```{r}
df3 <- t %>%
  select(all_of(levels3)) %>%
  rename(any_of(setNames(names(labels), labels)))
plot(likert(df3))
```

```{r}
df5 <- t %>%
  select(all_of(levels5)) %>%
  rename(any_of(setNames(names(labels), labels)))
plot(likert(df5))
```

>>>>

---
title: "Remote Collection of Respiratory Samples"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r libraries}
#| include: false
library(tidyverse)
library(lubridate)
library(gtsummary)
library(irr)
library(eulerr)
library(kableExtra)
library(caret)
library(likert)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning, include=FALSE}
# Use REDCap import
source(paste0(github_dir, "/CF/Jordana Hoppe/Home Collection/home_collection_data.r"))
source(paste0(github_dir, "/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r"))
# Exclude missing subjects
data_exclude <- c("009-07", "009-17", "009-25", "009-37")
survey_exclude <- c(15, 27, 30, 34, 47)
# Fix missing labels
labels <- label(data)[which(label(data) != "")]
labels[paste0(names(labels), ".factor")] <- as.character(labels)
survey_labels <- label(survey_data)[which(label(survey_data) != "")]
survey_labels[paste0(names(survey_labels), ".factor")] <- as.character(survey_labels)
labels <- c(labels, survey_labels)
# Combine race columns
races <- c(
  "American Indian or Alaska Native", "Asian", "Black or African-American",
  "Native Hawaiian or Other Pacific Islander", "White", "More than one race",
  "Unknown or not reported"
)
data$race <- apply(data, 1, function(r) {
  w <- which(r[paste0("race___", 1:7)] == 1)
  paste0(races[w], collapse = "/")
})
# Age
data$age <- round(as.numeric(ymd(data$clinic_date) - ymd(data$pt_dob)) / 365.25, 1)
# Find tests with no growth
data$clinic_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("clinic_cx_results___\\d$", names(r))]), na.rm = T)
})
data$clinic_cx_no_growth.factor <- factor(data$clinic_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
data$res_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("res_cx_results___\\d$", names(r))]), na.rm = T)
})
data$res_cx_no_growth.factor <- factor(data$res_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
data$home_cx_no_growth.factor <- apply(data, 1, function(r) {
  sum(as.numeric(r[grep("home_cx_results___\\d$", names(r))]), na.rm = T)
})
data$home_cx_no_growth.factor <- factor(data$home_cx_no_growth.factor,
  levels = 0:2,
  labels = c("No Growth", "Some Growth", "Some Growth")
)
# Additional labels for tables
labels["race"] <- "Race"
labels["age"] <- "Age at Clinic Visit"
# Genotype format
data$Genotype <- paste0(
  data$mutation1.factor == "F508del", "/", data$mutation2.factor == "F508del"
)
data$Genotype <- factor(data$Genotype,
  levels = c("TRUE/TRUE", "TRUE/FALSE", "FALSE/FALSE"),
  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
```

@@

## Table 3: Immediate Culture Results

```{r}
#| label: tbl-3
#| tbl-cap: Immediate Culture Results
data %>%
  filter(!subject_id %in% data_exclude) %>%
  select(clinic_cx_source.factor:clinic_burk_sp.factor) %>%
  tbl_summary(
    label = labels, missing_text = "Missing"
  ) %>%
  as_kable_extra()
```

@@

```{r}
agreement <- function(result_pattern, df = data) {
  cols <- paste0(
    c("clinic_cx_", "res_cx_", "home_cx_"),
    paste0(result_pattern, ".factor")
  )
  d <- df[, cols]
  colnames(d) <- c("Immediate", "Delayed", "Home")
  d <- d %>%
    mutate(across(everything(), ~ as.character(.)))
  d[d == "NA"] <- NA
  d[d == "Unchecked"] <- "Absent"
  d[d == "Checked"] <- "Present"
  fk <- kappam.fleiss(d)
  cat("\n")
  cat("### Agreement between all three methods")
  cat("\n")
  cat("\n")
  cat("#### Fleiss' kappa")
  cat("\n")
  cat("\n")
  f <- ftable(d)
  f %>%
    as.data.frame(.) %>%
    kable(.) %>%
    kable_styling(., "striped") %>%
    print(.)
  cat("\n")
  cat("\n")
  cat(
    "Overall Fleiss' kappa for the three tests was", round(fk$value, 3),
    paste0("(p = ", format.pval(fk$p.value, digits = 3, eps = 0.001), ")")
  )
  cat("\n")
  cat("\n")
  cat("#### Euler diagram")
  cat("\n")
  d %>%
    mutate(across(everything(), ~ as.numeric(. == "Present" | . == "No Growth"))) %>%
    drop_na(.) %>%
    euler(.) %>%
    plot(., quantities = T) %>%
    print(.)
  cat("\n")
  cat("\n")
  cat("### Pairwise agreement")
  cat("\n")
  d <- data.frame(lapply(d, factor, levels = c("Absent", "Present")))
  cat("\n")
  cat("#### Immediate vs. delayed (immediate as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Delayed"]),
    reference = d[, "Immediate"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("#### Immediate vs. home (immediate as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Home"]),
    reference = d[, "Immediate"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("#### Home vs. delayed (delayed as reference)")
  cat("\n")
  cat("\n")
  cat("##### Confusion matrix")
  cat("\n")
  cat("\n")
  c <- confusionMatrix((d[, "Home"]),
    reference = d[, "Delayed"], positive = "Present", mode = "everything"
  )
  print(kable(c$table, digits = 3))
  cat("\n")
  cat("\n")
  cat("##### Agreement")
  cat("\n")
  cat("\n")
  print(kable(c$overall, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
  cat("##### Sensitivity, specificity, etc.")
  cat("\n")
  cat("\n")
  print(kable(c$byClass, digits = 3, col.names = NULL))
  cat("\n")
  cat("\n")
}
```

@@

```{r}
yn_df <- t %>%
  select(all_of(yn)) %>%
  rename(any_of(setNames(names(labels), labels)))
yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")), check.names = F)
p <- plot(likert(yn_df))
ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/yn.png",
  plot = p,
  width = 8, height = 6, units = "in"
)
```

```{r}
check_df <- t %>%
  select(all_of(checkboxes)) %>%
  rename(any_of(setNames(names(labels), labels)))
colnames(check_df) <- c(
  "Uncooperative child", "Written instructions for collection were not clear",
  "I did not think the sample I obtained was a good sample", "Other",
  "Decreased travel time", "Less missed work",
  "Financial reasons - costs to travel to clinic can be a burden",
  "Health concerns with coming to clinic", "Other",
  "I do not feel comfortable with home collection",
  "I do not feel collection at home is reliable",
  "We prefer in-clinic visits only", "Other"
)
benefits_df <- check_df[, c(
  "Decreased travel time", "Less missed work",
  "Financial reasons - costs to travel to clinic can be a burden",
  "Health concerns with coming to clinic"
)]
barriers_df <- check_df[, c(
  "Uncooperative child",
  "I did not think the sample I obtained was a good sample",
  "Written instructions for collection were not clear",
  "I do not feel collection at home is reliable",
  "I do not feel comfortable with home collection"
)]

p_benefits <- plot(likert(benefits_df)) + ggtitle("Benefits") +
  theme(
    axis.text.x = element_blank(), axis.title.x = element_blank(),
    legend.position = "none"
  )
p_barriers <- plot(likert(barriers_df)) + ggtitle("Barriers")
p <- p_benefits / p_barriers
ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/check.png",
  plot = p,
  width = 8, height = 6, units = "in"
)
```

```{r}
df3 <- t %>%
  select(all_of(levels3)) %>%
  rename(any_of(setNames(names(labels), labels)))
p <- plot(likert(df3))
ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/q2.png",
  plot = p,
  width = 8, height = 6, units = "in"
)
```

```{r}
df5 <- t %>%
  select(all_of(levels5)) %>%
  rename(any_of(setNames(names(labels), labels)))
p <- plot(likert(df5))
ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/difficulty.png",
  plot = p,
  width = 9, height = 6, units = "in"
)
```
